---
title: "Bellabeat_Capstone"
author: "Nanke Williams"
date: "2022-09-29"
output: html_document
---

### Introduction and Background

Bellabeat has 5 products: an app (Bellabeat app), a wellness tracker, a smart watch, a water bottle line and a subscription membership service. 

The business task is to analyze user smart device data collected by the wearable technology and tracked on the app, in order to:

* gain insight into how customers are using their smart devices, and
* recommend appropriate marketing strategies for the company.

### Sourcing the Data
The Bellabeat dataset was sourced from [here](https://www.kaggle.com/arashnic/fitbit). The primary tables used in this analysis are:

* dailyActivity_merged.csv
* sleepDay_merged.csv
* weightLogInfo_merged.csv
* heartrate_records.csv

### Analysing the Credibility of the dataset
The data contains the following number of unique user records: 

* dailyActivity_merged.csv - 33 records
* sleepDay_merged.csv - 24 records
* weightLogInfo_merged.csv - 8 records
* heartrate_records.csv - 14 records

There's a very limited amount of data to analyze and make recommendations from. The data is accurate, original and comprehensive. But it is dated and not current.

### Preliminary analysis of the datasets
Preliminary analysis of the dataset was conducted in SQL. The following points were found:

* There are only 3 unique IDs identical to all 4 datasets. This sample is too small to analyze. We have chosen instead to prioritize the dailyActivity dataset and join relevant variables from the other datasets to it. A new summary dataset was created.
* There are null records across all variables in the largest dataset (dailyActivity)
* This will affect all cross data analysis and significantly reduce the sample size. 

In R, we will perform correlation analysis and visualizations on some on the relationship between some of the variables. 

### Installing and loading common packages and libraries
```{r}
# install.packages('tidyverse')
library(tidyverse)
```

### Loading your CSV files
We create variables and load the csv files into them.
```{r}
daily_activity <- read.csv("dailyActivity_merged.csv")
sleep_log <- read.csv("sleepDay_merged.csv")
weight_log <- read.csv("weightLogInfo_merged.csv")
heartrate_log <- read.csv("heartrate_records.csv")
summary_table <- read.csv('fitabasecollated_table.csv')
```

### Preliminary analysis
We conduct preliminary analysis of the tables using head(), glimpse(), str(), glimpse(), summary() methods to get an initial overview of the data.

```{r}
head(daily_activity)
```

```{r}
head(sleep_log)
```

```{r}
head(weight_log)
```

```{r}
head(heartrate_log)
```

```{r}
head(summary_table)
```
This is the primary table to be used for this review.

str() shows us the column names, data types and a sample of the records in each variable.
```{r}
str(summary_table)
```

## Data Cleaning
For proper analysis, we will convert the following columns to appropriate datatypes: 

* id column to character

```{r}
summary_table$id <- as.character(summary_table$id)
```

A glimpse of the table after cleaning.
```{r}
glimpse(summary_table)
```
Everything looks ready for analysis.

### Summary statistics
First, some summary statistics to view statistical information on all variables.
```{r}
summary(summary_table)
```
Some variables (such as avg_weight_kg,avg_bmi, sleep_records, mins_asleep, mins_in_bed contain significant null values which will affect visual analysis)

A brief summary of select columns from each table (daily_activity, sleepday and weightlog_info), shows the following:
Daily Activity Table - Steps
```{r}
summary(summary_table$steps)
```

Sleep Log Table - minutes asleep
```{r}
summary(summary_table$mins_asleep)
```

Weight Log Table - average weight in kilograms
```{r}
summary(summary_table$avg_weight_kg)
```

#### Visualizing the Data
First, we measure correlation between the relevant variables and then visualize that correlation in a scatter plot.

* A correlation coefficient is a metric that measures the extent to which numeric variables are associated with one another. 
* The correlation coefficient always lies between +1 (perfect positive correlation) and –1 (perfect negative correlation); 0 indicates no correlation.

Measure correlation between active distance and active minutes
```{r}
cor(x= summary_table$active_distance, y = summary_table$active_mins, use = 'complete.obs', method = 'pearson')
```
This shows significant association between the variables. Let's confirm with a scatterplot chart.

Chart relationship between active minutes and active distance
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = active_distance, y = active_mins)) +
  labs(title = 'Relationship between Active Minutes & Active Distance', 
       caption = 'Active Minutes & Distance = Total Mins - Sedentary Minutes')
```

Measure correlation betwen steps and active distance
```{r}
cor(x= summary_table$active_distance, y = summary_table$steps, use = 'everything', method = 'pearson')
```

Chart relationship between steps and active distance
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = active_distance, y = steps)) +
  labs(title = 'Relationship between Steps & Active Distance', 
       caption = 'There is a positive correlation between steps and active distance')
```

Measure correlation between calories and active minutes
```{r}
cor(x= summary_table$calories, y = summary_table$active_mins, use = 'everything', method = 'pearson')
```

Chart relationship between calories and active minutes
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = active_mins, y = calories)) +
  labs(title = 'Relationship between Calories & Active Minutes', 
       caption = 'Correlation between these variables not clear')
```
The relationship between these variables are not distint enough. More data is needed to confirm or exclude any conclusions about correlation.

Measure correlation between steps and average bmi
```{r}
cor(x= summary_table$steps, y = summary_table$avg_bmi, use = 'complete.obs', method = 'pearson')
```
There is a significant negative correlation between bmi and step activity. This will be clearer with a visual graph.

Chart relationship between steps and avg_bmi
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = steps, y = avg_bmi)) +
  labs(title = 'Relationship between Steps & Average Body Mass Index', 
       caption = 'Negative Correlation Exists. Sample too small')
```
Unfortunately, the number of records available for this chart (8 records) is too small to make any final conclusions.

Measure correlation between steps and average weight
```{r}
cor(x= summary_table$steps, y = summary_table$avg_weight_kg, use = 'complete.obs', method = 'pearson')
```

Chart relationship between steps and avg_weight
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = steps, y = avg_weight_kg)) +
    labs(title = 'Relationship between Steps & Average Weight', 
       caption = 'Negative Correlation Exists. Sample too small')
```

Measure correlation between steps and minutes asleep
```{r}
cor(x= summary_table$steps, y = summary_table$mins_asleep, use = 'complete.obs', method = 'pearson')
```
There is no significant correlation between active steps and minutes asleep. The visual chart will show this much clearly.

Chart relationship between steps and mins_asleep
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = steps, y = mins_asleep)) + 
    labs(title = 'Relationship between Steps & Minutes Asleep', 
       caption = 'No Correlation')
```

Measure correlation between calories and minutes asleep
```{r}
cor(x= summary_table$calories, y = summary_table$mins_asleep, use = 'complete.obs', method = 'pearson')
```

Chart relationship between calories and mins_asleep
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = calories, y = mins_asleep)) + 
      labs(title = 'Relationship between Calories & Minutes Asleep', 
       caption = 'No Correlation')
```

Measure correlation between minutes asleep and active minutes
```{r}
cor(x= summary_table$mins_asleep, y = summary_table$active_mins, use = 'complete.obs', method = 'pearson')
```

Chart relationship between minutes asleep and active minutes
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = mins_asleep, y = active_mins)) +
      labs(title = 'Relationship between Minutes Asleep & Active Minutes', 
       caption = 'Correlation is unclear')
```

Measure correlation between calories and average weight 
```{r}
cor(x= summary_table$calories, y = summary_table$avg_weight_kg, use = 'complete.obs', method = 'pearson')
```

Chart relationship between calories and average weight 
```{r}
ggplot(data = summary_table) + 
  geom_point(mapping = aes(x = calories, y = avg_weight_kg)) + 
      labs(title = 'Relationship between Calories & Average Weight', 
       caption = 'No Correlation')
```


#### Conclusion from Analysis
Here are some recommendations from the analysis of the dataset:

* Bellabeat should provide more data for analysis
* Bellabeat should market to its users via regular reminders on the Bellabeat app, that they wear their smart technology
* Although a small sample size, the negative correlation between step activity and average body mass index and average weight is a positive metric. This should be highlighted to users
* Important correlations could not be properly determined because of the small sample size - e.g. the relationship between calories, activity, sleep and weight is not clear.

